---
title: "Prepare bam files"
output:
    BiocStyle::html_document:
        toc: true
author: "Charles Joly Beauparlant"
date: "September 30, 2015"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
\usepackage[utf8]{inputenc}
---

```{r devtools, message=FALSE}
# TODO: Replace with library(metageneInputs) when available.
#library(devtools)
#devtools::load_all()
library(metageneInputs)
```

## Introduction

The goals of this vignette are:

* To prepare the `designs.RData` object (which contains `designs_encode` and
`designs_validation`.
* To produce the script (Makefile) for downloading all the bam files for the
metagene paper.

## Designs

### ENCODE

The main part of the analysis consists in producing the metagene objects for the
ENCODE dataset for the GM12878 cell line. In order to do so, we must first fetch
the names of the targets which have bam files available in ENCODE:

```{r fetchTargets}
cell_line <- "GM12878"
targets <- get_encode_targets(cell_line)
targets
```

We need to remove the histones:
```{r removeHistones}
i <- !grepl("H[0-9]K", targets)
targets <- targets[i]
targets
```

We need to remove the control:
```{r removeControl}
i <- targets != "Control" & targets != "mouse" & targets != "rabbit"
targets <- targets[i]
targets
```

We produce a design for each target:
```{r produceDesigns, cache=TRUE}
# We use the log hack to avoid all the outpus from the queryEncode function
log <- capture.output({
    designs_encode <- lapply(targets, get_encode_design, cell_line)
})
names(designs_encode) <- targets
head(designs_encode)
```

### Validation

The validation of the results will be performed on the
["Variation and genetic control of chromatin architecture in humans"](https://www.ebi.ac.uk/arrayexpress/experiments/E-MTAB-3657/files/)
dataset.

#### Experiment descriptions

Load and clean the descriptions:

```{r loadDescriptions}
library(tidyr)
library(dplyr)
file_descriptions <- system.file("extdata/E-MTAB-3657.sdrf.txt",
                                package = "metageneInputs")
descriptions <- read.table(file_descriptions, header = TRUE, sep = "\t",
                           stringsAsFactors = FALSE)
to_keep <- c("Source.Name", "Comment..Derived.ArrayExpress.FTP.file.")
descriptions <- descriptions[,colnames(descriptions) %in% to_keep]
colnames(descriptions) <- c("individual_target", "url")
descriptions <- tbl_df(descriptions) %>%
                    separate(individual_target,
                             into = c("individual", "target"), sep = "_") %>%
                    filter(target %in% c("PU.1", "RPB2"))
```

Create the design:

```{r designValidation}
design_validation <- data.frame(bam_files = basename(descriptions$url))
i <- descriptions[["target"]] == "PU.1"
j <- descriptions[["target"]] == "RPB2"
design_validation[["PU.1"]] <- 0
design_validation[["RPB2"]] <- 0
design_validation[["PU.1"]][i] <- 1
design_validation[["RPB2"]][j] <- 1
design_validation
```

Save the results:

```{r save}
save(designs_encode, design_validation, file = "designs.RData")
```

## Makefile

We will now produce the Makefile to download the bam files.

```{r}
fileConn<-file("Makefile", open = "w")
writeLines("# Targets", fileConn)
close(fileConn)
fileConn<-file("Makefile", open = "a")
writeLines("## ENCODE's files", fileConn)
bam_files <- unique(as.character(unlist(lapply(designs_encode,
                                               `[[`, "bam_files"))))
writeLines(paste0("TARGETS_ENCODE+=", bam_files), fileConn)
writeLines("", fileConn)
writeLines("## Validation's files", fileConn)
bam_files <- as.character(design_validation[["bam_files"]])
writeLines(paste0("TARGETS_VALIDATION+=", bam_files), fileConn)
writeLines("", fileConn)
writeLines("# URLs", fileConn)
writeLines("URL_ENCODE=https://www.encodeproject.org/files", fileConn)
writeLines("URL_VALIDATION=ftp://ftp.ebi.ac.uk/pub/databases/microarray/data/experiment/MTAB/E-MTAB-3657",
           fileConn)
writeLines("", fileConn)
writeLines("# Phony targets", fileConn)
writeLines(".PHONY: all", fileConn)
writeLines("all: $(TARGETS_ENCODE) $(TARGETS_VALIDATION)", fileConn)
writeLines("", fileConn)
writeLines("# Recipes", fileConn)
writeLines("ENCFF%.bam:", fileConn)
writeLines("\twget $(URL_ENCODE)/ENCFF$*/@@download/$@", fileConn)
writeLines("", fileConn)
writeLines("E-MTAB%.bam:", fileConn)
writeLines("\twget $(URL_VALIDATION)/$@", fileConn)
close(fileConn)
```
