---
title: "Split regions"
output:
    BiocStyle::html_document:
        toc: true
author: "Charles Joly Beauparlant"
date: "September 29, 2015"
vignette: >
  %\VignetteIndexEntry{Split regions.}
  %\VignetteEngine{knitr::rmarkdown}
\usepackage[utf8]{inputenc}
---

```{r devtools, message=FALSE}
library(metageneInputs)
library(ENCODExplorer)
```

## Introduction

The goal of this vignette is to produce the groups of regions that will be used
for the metagene paper.

All the files produced will be save inside a temporary directory and will need
to be moved manually:
```{r outputDir}
outdir <- basename(tempfile(""))
dir.create(outdir)
outdir
```

## Regions

To obtain the robust regoins, we need to keep only the regions that overlap
with the TSS from the `FantomTSS.hg19` package and then extract the mean TPM
value for each regions.

### Robust enhancer regions

```{r robustEnhancerRegions}
cell_line <- "GM12878"
robust_enhancer_regions <- get_robust_enhancer_regions(cell_line,
                                                        keep_y = FALSE)
robust_enhancer_regions
```

### Robust promoter regions

```{r robustPromoterRegions}
robust_promoter_regions <- get_robust_promoter_regions(cell_line,
                                                        keep_y = FALSE)
robust_promoter_regions
```

## Tresholds

In order to split promoters and enhancers into 5 different groups based on
their expression level in TPM, we will first calculate tresholds for each
groups using every TPM values from every non-zero robust TSS TPM of Fantom5
except those from the `r cell_line`.

```{r tresholds, cache=TRUE}
fantom_tpm <- get_all_tss_tpm(exclude = cell_line)
probs <- seq(1.0, 0, -0.25)
quantiles <- quantile(fantom_tpm[fantom_tpm != 0], probs = probs)
quantiles
```

## Split the regions

The enhancers:

```{r splitEnhancers}
enhancers <- split_by_quantiles(robust_enhancer_regions, quantile, cell_line)
enhancers
```

The promoters:

```{r splitPromoters}
promoters <- split_by_quantiles(robust_promoter_regions, quantile, cell_line)
promoters
```

## Save results:

```{r save}
outfile <- paste0(outdir, "/splitted_regions.RData")
save(enhancers, promoters, file = outfile)
```

## TPM `GRangesList`

```{r tpmGRangesList}
GenomicRanges::mcols(robust_promoter_regions)[["gene_id"]] <- NULL
TPM <- GenomicRanges::GRangesList(enhancers = robust_enhancer_regions,
                                  promoters = robust_promoter_regions)
outfile <- paste0(outdir, "/tpm_grangeslist.RData")
save(TPM, file = outfile)
```
