---
title: "Split regions"
author: "Charles Joly Beauparlant"
date: "September 29, 2015"
output: html_document
---

```{r devtools, warning=FALSE, message=FALSE}
# TODO: Replace with library(metageneInputs) when available.
library(devtools)
devtools::load_all()
```

## Introduction

The goal of this vignette is to produce the groups of regions that will be used
for the metagene paper.

## Regions

To obtain the robust regoins, we need to keep only the regions that overlap
with the TSS from the `FantomTSS.hg19` package and then extract the mean TPM
value for each regions.

### Robust enhancer regions

```{r robustEnhancerRegions}
cell_line <- "GM12878"
robust_enhancer_regions <- get_robust_enhancer_regions(cell_line,
                                                        keep_y = FALSE)
robust_enhancer_regions
```

### Robust promoter regions

```{r robustPromoterRegions}
robust_promoter_regions <- get_robust_promoter_regions(cell_line,
                                                        keep_y = FALSE)
robust_promoter_regions
```

## Tresholds

In order to split promoters and enhancers into 5 different groups based on
their expression level in TPM, we will first calculate tresholds for each
groups using every TPM values from every non-zero robust TSS TPM of Fantom5
except those from the `r cell_line`.

```{r tresholds, cache=TRUE}
fantom_tpm <- get_all_tss_tpm(exclude = cell_line)
probs <- seq(1.0, 0, -0.25)
quantiles <- quantile(fantom_tpm[fantom_tpm != 0], probs = probs)
quantiles
```

## Split the regions

The enhancers:

```{r splitEnhancers}
enhancers <- split_by_quantiles(robust_enhancer_regions, quantile, cell_line)
enhancers
```

The promoters:

```{r splitPromoters}
promoters <- split_by_quantiles(robust_promoter_regions, quantile, cell_line)
promoters
```

## Save results:

```{r save}
save(enhancers, promoters, file = "splitted_regions.RData")
```
